### 2025-01-29 reports/rmd-report-template.Rmd
Rmd report template fixed.

---
title: "RMD Report Template"
author: "John Doe"
date: "`r Sys.Date()`"
output:
  html_document: 
    fig_caption: true
  pdf_document: 
    df_print: paged
    toc: true
    fig_caption: true
    keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.