### 2025-02-03 reports/capstone-movielens-report.pdf-template.Rmd
PDF template added for Capstone MovieLens report.

---
title: "Capstone Movielens Report"
author: "Azamat Kurbanaev"
date: "`r Sys.Date()`"
bibliography: references.bib
#csl: bit-numerical-mathematics.csl 
output:
  pdf_document: 
    df_print: paged
    toc: true
    fig_caption: true
    keep_tex: true
    citation_package: biblatex
  # html_document: 
  #   fig_caption: true
colorlinks: true
linkcolor: red
citecolor: blue
header-includes:
    - \usepackage{hyperref}
    - \usepackage[natbib=true, 
                  style=numeric, 
                  backref=true, 
                  sorting=none]{biblatex}
    - \hypersetup{backref,
          pdfpagemode=FullScreen,
          colorlinks=true,
          implicit=false}
vignette: >
  %\VignetteIndexEntry{RMarkdown Citations - Numeric Style}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::knitr}
---

```{r setup, include=TRUE, echo = FALSE}

knitr::opts_chunk$set(echo = TRUE)
```

## Introduction / Overview / Executive Summary

The goal of the project is to build a Recommendation System using a [10M version of the MovieLens dataset](http://grouplens.org/datasets/movielens/10m/).
Following the _Netflix Grand Prize Contest_[@NYT1MNP] requirements, we will evaluate the RMSE score defined as:

$$
\mbox{RMSE} = \sqrt{\frac{1}{N} \sum_{i,j}^{N} (y_{i,j} - \hat{y}_{i,j})^2}
$$

with $N$ being the number of user/movie combinations for which we make predictions and the sum occurring over all these combinations[@IDS2_23-2].

Our goal is to achieve a value of less than 0.86490 (compare with the _Netflix Grand Prize_ requirement: of at least 0.8563[@BigChaosSln]).

### Dataset

```{r cars}
summary(cars)
```

## Methods / Analysis

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Conclusion

Hello Conclusion!

This is a great conclusion, isn't it?!!