### 2025-02-03 reports/capstone-movielens-report.html-template.Rmd
HTML template added for Capstone MovieLens report.

---
title: "Capstone Movielens Report"
author: "Azamat Kurbanaev"
date: "`r Sys.Date()`"
output:
  html_document: 
    fig_caption: true
vignette: >
  %\VignetteIndexEntry{RMarkdown Citations - Alphabetic Style}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=TRUE, echo = FALSE}
library(RefManageR)
bib <- ReadBib("references.bib", check = FALSE)
BibOptions(check.entries = FALSE, style = "markdown", bib.style = "numeric", cite.style = 'numeric')

citeB <- function(citeKey){
  AutoCite(bib, citeKey, .opts = list(hyperlink = "to.bib", super = FALSE))
}

knitr::opts_chunk$set(echo = TRUE)
```

## Introduction / Overview / Executive Summary

The goal of the project is to build a Recommendation System using a [10M version of the MovieLens dataset](http://grouplens.org/datasets/movielens/10m/).
Following the [Netflix Grand Prize Contest](https://archive.nytimes.com/bits.blogs.nytimes.com/2009/09/21/netflix-awards-1-million-prize-and-starts-a-new-contest/index.html) requirements, we will evaluate the RMSE score defined as:

$$
\mbox{RMSE} = \sqrt{\frac{1}{N} \sum_{i,j}^{N} (y_{i,j} - \hat{y}_{i,j})^2}
$$

with $N$ being the number of user/movie combinations for which we make predictions and the sum occurring over all these combinations`r citeB("IDS2_23-2")`.

Our goal is to achieve a value of less than 0.86490 (compare with the _Netflix Grand Prize_ requirement: of at least 0.8563`r citeB("BigChaosSln")`).

### Dataset

```{r cars}
summary(cars)
```

## Methods / Analysis

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Conclusion

Hello Conclusion!

This is a great conclusion, isn't is?!!

**References**

```{r results = "asis", echo = FALSE}
PrintBibliography(bib, .opts = list(check.entries = FALSE))
```