### 2025-02-07 reports/capstone-movielens-report.Rmd
Report: `edx` dataset analisys updated (additional data visualization code added).

---
title: "Capstone Movielens Report"
author: "Azamat Kurbanaev"
date: "`r Sys.Date()`"
bibliography: references.bib
#csl: bit-numerical-mathematics.csl 
output:
  pdf_document: 
    df_print: paged
    toc: true
    fig_caption: true
    keep_tex: true
    citation_package: biblatex
colorlinks: true
linkcolor: red
citecolor: blue
header-includes:
    - \usepackage{hyperref}
    - \usepackage[natbib=true, 
                  style=numeric, 
                  backref=true, 
                  sorting=none]{biblatex}
    - \hypersetup{backref,
          pdfpagemode=Normal,
          colorlinks=true,
          implicit=false}
vignette: >
  %\VignetteIndexEntry{RMarkdown Citations - Numeric Style}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::knitr}
---

```{r setup, include=TRUE, echo = FALSE}
# if(!require(tidyverse)) 
#   install.packages("tidyverse", repos = "http://cran.us.r-project.org")
# if(!require(caret)) 
#   install.packages("caret", repos = "http://cran.us.r-project.org")
# if(!require(data.table)) 
#   install.packages("data.table", repos = "http://cran.us.r-project.org")
if(!require(pak)) 
  install.packages("pak")

# Loading the required libraries
library(caret)
library(cowplot)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(lubridate)
library(Metrics)
library(recosystem)
library(scales)
library(stringr)
library(tibble)
library(tidyr)
library(pak)

knitr::opts_chunk$set(echo = TRUE)
```

## Introduction / Overview / Executive Summary

The goal of the project is to build a Recommendation System using a [10M version of the MovieLens dataset](http://grouplens.org/datasets/movielens/10m/).
Following the [Netflix Grand Prize Contest](https://archive.nytimes.com/bits.blogs.nytimes.com/2009/09/21/netflix-awards-1-million-prize-and-starts-a-new-contest/index.html) requirements, we will evaluate the _Root Mean Square Error_ (_RMSE_) score defined as:

$$
\mbox{RMSE} = \sqrt{\frac{1}{N} \sum_{i,j}^{N} (y_{i,j} - \hat{y}_{i,j})^2}
$$

with $N$ being the number of user/movie combinations for which we make predictions and the sum occurring over all these combinations[@IDS2_23-2].

Our goal is to achieve a value of less than 0.86490 (compare with the _Netflix Grand Prize_ requirement: of at least 0.8563[@BigChaosSln]).

### Datasets

``` {r , echo = FALSE}
```

To start with we have to generate two datasets derived from the _MovieLens_ one mentioned above:

* **edx:** we use it to develop and train our algorithms;
* **final_holdout_test:**  according to the course requirements, we use it exclusively to evaluate the _**RMSE**_ of our final algorithm.

For this purpose the following package has been developed by the author of this report: `edx.capstone.movielens.data`. The source code of the package is available [on GitHub](https://github.com/AzKurban-edX-DS/edx.capstone.movielens.data)[@edx_capstone_movielens_data].

Let us install the development version of this package from the GitHub repository and attach the correspondent library to the global environment:
```{r }
if(!require(edx.capstone.movielens.data)) pak::pak("AzKurban-edX-DS/edx.capstone.movielens.data")

library(edx.capstone.movielens.data)
edx <- edx.capstone.movielens.data::edx
final_holdout_test <- edx.capstone.movielens.data::final_holdout_test

summary(edx)
summary(final_holdout_test)
```

Let us look into the details of the `edx` dataset:
``` {r echo = TRUE}
str(edx)
```
Note that we have 9000055 rows and six columns in there:  
```{r }
dim(edx)
```

Also, we can see that no movies have a rating of 0. Movies are rated from 0.5 to 5.0 in 0.5 increments:
```{r }
#library(dplyr)
s <- edx |> group_by(rating) |>
  summarise(n = n())
print(s)
```

Next, we have 10677 different movies 
```{r}
n_distinct(edx$movieId)
```
and 69878 different users in the dataset:
```{r}
n_distinct(edx$userId)
```

The following code computes movie rating summaries by popular genres like Drama, Comedy, Thriller, and Romance:
```{r}
#library(stringr)
genres = c("Drama", "Comedy", "Thriller", "Romance")
sapply(genres, function(g) {
  sum(str_detect(edx$genres, g))
})
```

Further, we can find out the movies that have the greatest number of ratings using the following code:
```{r}
ordered_movie_ratings <- edx |> group_by(movieId, title) |>
  summarize(number_of_ratings = n()) |>
  arrange(desc(number_of_ratings))
print(head(ordered_movie_ratings))
```

and figure out the most given ratings in order from most to least:
```{r}
ratings <- edx |>  group_by(rating) |>
     summarise(count = n()) |>
     arrange(desc(count))
print(ratings)
```

The following code allows us to summarize that in general, half-star ratings are less common than whole-star ratings (e.g., there are fewer ratings of 3.5 than there are ratings of 3 or 4, etc.):
```{r}
print(edx |> group_by(rating) |> summarize(count = n()))
```

We can visually see that from the following plot:
```{r}
edx |>
  group_by(rating) |>
  summarize(count = n()) |>
  ggplot(aes(x = rating, y = count)) +
  geom_line()

```

Further analysis of the `edx` dataset was inspired by the article: A Movie Recommender System Using R - BEST[@MRS-R-BEST], from which the code and explanatory notes below were cited.

#### Rating distribution plot[@MRS-R-BEST]
\
The code below demonstrates another way of visualizing the rating distribution:
```{r}
edx |>
  group_by(rating) |>
  summarize(count = n()) |>
  ggplot(aes(x = rating, y = count)) +
  geom_bar(stat = "identity", fill = "#8888ff") +
  ggtitle("Rating Distribution") +
  xlab("Rating") +
  ylab("Occurrences Count") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(n.breaks = 10) +
  theme_economist() +
  theme(axis.title.x = element_text(vjust = -5, face = "bold"), 
        axis.title.y = element_text(vjust = 10, face = "bold"), 
        plot.margin = margin(0.7, 0.5, 1, 1.2, "cm"))

```

This graph is another confirmation of what we found out above: rounded ratings occur more often than half-stared ones. The upward trend previously discussed is now perfectly clear, although it seems to top right between the 3 and 4-star ratings lowering the occurrences count afterward. That might be due to users being more hesitant to rate with the highest mark for whichever reasons they might hold[@MRS-R-BEST].

#### Yearly rating count[@MRS-R-BEST]
\
```{r}
print(edx |> 
  mutate(year = year(as_datetime(timestamp, origin = "1970-01-01"))) |>
  group_by(year) |>
  summarize(count = n())
)
```

#### Average rating per year plot[@MRS-R-BEST]
\
```{r}
edx |> 
  mutate(year = year(as_datetime(timestamp, origin = "1970-01-01"))) |>
  group_by(year) |>
  summarize(avg = mean(rating)) |>
  ggplot(aes(x = year, y = avg)) +
  geom_bar(stat = "identity", fill = "#8888ff") + 
  ggtitle("Average rating per year") +
  xlab("Year") +
  ylab("Average rating") +
  scale_y_continuous(labels = comma) + 
  theme_economist() +
  theme(axis.title.x = element_text(vjust = -5, face = "bold"), 
        axis.title.y = element_text(vjust = 10, face = "bold"), 
        plot.margin = margin(0.7, 0.5, 1, 1.2, "cm"))
```

#### Movie popularity count[@MRS-R-BEST]
\
```{r}
print(edx |> 
  group_by(movieId) |> 
  summarize(count = n()) |>
  slice_head(n = 10)
)
```

```{r}
summary(edx |> group_by(movieId) |> summarize(count = n()) |> select(count))
```

#### Ratings per movie plot[@MRS-R-BEST]
\
```{r}
edx |>
  group_by(movieId) |>
  summarize(count = n()) |>
  ggplot(aes(x = movieId, y = count)) +
  geom_point(alpha = 0.2, color = "#4020dd") +
  geom_smooth(color = "red") +
  ggtitle("Ratings per movie") +
  xlab("Movies") +
  ylab("Number of ratings") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(n.breaks = 10) +
  theme_economist() +
  theme(axis.title.x = element_text(vjust = -5, face = "bold"), 
        axis.title.y = element_text(vjust = 10, face = "bold"), 
        plot.margin = margin(0.7, 0.5, 1, 1.2, "cm"))
```

#### Movies' rating histogram[@MRS-R-BEST]
\
```{r}
edx |>
  group_by(movieId) |>
  summarize(count = n()) |>
  ggplot(aes(x = count)) +
  geom_histogram(fill = "#8888ff", color = "#4020dd") +
  ggtitle("Movies' rating histogram") +
  xlab("Rating count") +
  ylab("Number of movies") +
  scale_y_continuous(labels = comma) +
  scale_x_log10(n.breaks = 10) +
  theme_economist() +
  theme(axis.title.x = element_text(vjust = -5, face = "bold"), 
        axis.title.y = element_text(vjust = 10, face = "bold"), 
        plot.margin = margin(0.7, 0.5, 1, 1.2, "cm"))
```

#### Ratings per user[@MRS-R-BEST]

##### User rating count (activity measure)
\
```{r}
print(edx |> 
  group_by(userId) |> 
  summarize(count = n()) |>
  slice_head(n = 10)
)
```

##### User rating summary
\
```{r}
summary(edx |> group_by(userId) |> summarize(count = n()) |> select(count))
```

##### Ratings per user plot
\
```{r}
edx |>
  group_by(userId) |>
  summarize(count = n()) |>
  ggplot(aes(x = userId, y = count)) +
  geom_point(alpha = 0.2, color = "#4020dd") +
  geom_smooth(color = "red") +
  ggtitle("Ratings per user") +
  xlab("Users") +
  ylab("Number of ratings") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(n.breaks = 10) +
  theme_economist() +
  theme(axis.title.x = element_text(vjust = -5, face = "bold"), 
        axis.title.y = element_text(vjust = 10, face = "bold"), 
        plot.margin = margin(0.7, 0.5, 1, 1.2, "cm"))
```

##### Users' rating histogram
\
```{r}
edx |>
  group_by(userId) |>
  summarize(count = n()) |>
  ggplot(aes(x = count)) +
  geom_histogram(fill = "#8888ff", color = "#4020dd") +
  ggtitle("Users' rating histogram") +
  xlab("Rating count") +
  ylab("Number of users") +
  scale_y_continuous(labels = comma) +
  scale_x_log10(n.breaks = 10) +
  theme_economist() +
  theme(axis.title.x = element_text(vjust = -5, face = "bold"), 
        axis.title.y = element_text(vjust = 10, face = "bold"), 
        plot.margin = margin(0.7, 0.5, 1, 1.2, "cm"))
```

##### 
\
```{r}

```

##### 
\
```{r}

```

##### 
\
```{r}

```

##### 
\
```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

#### [@MRS-R-BEST]
\
```{r}

```

## Methods / Analysis

You can also embed plots, for example:

```{r pressure, echo=FALSE}
#plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Conclusion

Hello Conclusion!

This is a great conclusion, isn't it?!!