### 2025-02-13 reports/capstone-movielens-report.Rmd
Report: Movie effects & Penalized least squares described.

---
title: "Capstone Movielens Report"
author: "Azamat Kurbanaev"
date: "`r Sys.Date()`"
bibliography: references.bib
#csl: bit-numerical-mathematics.csl 
output:
  pdf_document: 
    df_print: paged
    toc: true
    fig_caption: true
    keep_tex: true
    citation_package: biblatex
colorlinks: true
linkcolor: red
citecolor: blue
header-includes:
    - \usepackage{hyperref}
    - \usepackage[natbib=true, 
                  style=numeric, 
                  backref=true, 
                  sorting=none]{biblatex}
    - \hypersetup{backref,
          pdfpagemode=Normal,
          colorlinks=true,
          implicit=false}
vignette: >
  %\VignetteIndexEntry{RMarkdown Citations - Numeric Style}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::knitr}
---

```{r setup, include=TRUE, echo = FALSE}
# if(!require(tidyverse)) 
#   install.packages("tidyverse", repos = "http://cran.us.r-project.org")
# if(!require(caret)) 
#   install.packages("caret", repos = "http://cran.us.r-project.org")
# if(!require(data.table)) 
#   install.packages("data.table", repos = "http://cran.us.r-project.org")
if(!require(pak)) 
  install.packages("pak")

# Loading the required libraries
library(caret)
library(cowplot)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(lubridate)
library(Metrics)
library(recosystem)
library(scales)
library(stringr)
library(tibble)
library(tidyr)

library(rafalib)
library(pak)

knitr::opts_chunk$set(echo = TRUE)
```

## Introduction / Overview / Executive Summary

The goal of the project is to build a Recommendation System using a [10M version of the MovieLens dataset](http://grouplens.org/datasets/movielens/10m/).
Following the [Netflix Grand Prize Contest](https://archive.nytimes.com/bits.blogs.nytimes.com/2009/09/21/netflix-awards-1-million-prize-and-starts-a-new-contest/index.html) requirements, we will evaluate the _Root Mean Square Error_ (_RMSE_) score, which, as shown in [Section 23.2 Loss function](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#sec-netflix-loss-function) of the _Course Textbook_, is defined as:
$$
\mbox{RMSE} = \sqrt{\frac{1}{N} \sum_{i,j}^{N} (y_{i,j} - \hat{y}_{i,j})^2}
$$

with $N$ being the number of user/movie combinations for which we make predictions and the sum occurring over all these combinations[@IDS2_23-2].

Our goal is to achieve a value of less than 0.86490 (compare with the _Netflix Grand Prize_ requirement: of at least 0.8563[@BigChaosSln]).

### Datasets

``` {r , echo = FALSE}
```

To start with we have to generate two datasets derived from the _MovieLens_ one mentioned above:

* **edx:** we use it to develop and train our algorithms;
* **final_holdout_test:**  according to the course requirements, we use it exclusively to evaluate the _**RMSE**_ of our final algorithm.

For this purpose the following package has been developed by the author of this report: `edx.capstone.movielens.data`. The source code of the package is available [on GitHub](https://github.com/AzKurban-edX-DS/edx.capstone.movielens.data)[@edx_capstone_movielens_data].

Let's install the development version of this package from the GitHub repository and attach the correspondent library to the global environment:
```{r }
if(!require(edx.capstone.movielens.data)) pak::pak("AzKurban-edX-DS/edx.capstone.movielens.data")

library(edx.capstone.movielens.data)
edx <- edx.capstone.movielens.data::edx
final_holdout_test <- edx.capstone.movielens.data::final_holdout_test

summary(edx)
summary(final_holdout_test)
```
#### `edx` Dataset
\
Let's look into the details of the `edx` dataset:
``` {r echo = TRUE}
str(edx)
```
Note that we have 9000055 rows and six columns in there:  
```{r }
dim_edx <- dim(edx)
print(dim_edx)
```

Also, we can see that no movies have a rating of 0. Movies are rated from 0.5 to 5.0 in 0.5 increments:
```{r }
#library(dplyr)
s <- edx |> group_by(rating) |>
  summarise(n = n())
print(s)
```

##### Movie Genres Data 
\
The following code computes movie rating summaries by popular genres like Drama, Comedy, Thriller, and Romance:
```{r}
#library(stringr)
genres = c("Drama", "Comedy", "Thriller", "Romance")
sapply(genres, function(g) {
  sum(str_detect(edx$genres, g))
})
```

Further, we can find out the movies that have the greatest number of ratings using the following code:
```{r}
ordered_movie_ratings <- edx |> group_by(movieId, title) |>
  summarize(number_of_ratings = n()) |>
  arrange(desc(number_of_ratings))
print(head(ordered_movie_ratings))
```

and figure out the most given ratings in order from most to least:
```{r}
ratings <- edx |>  group_by(rating) |>
     summarise(count = n()) |>
     arrange(desc(count))
print(ratings)
```

The following code allows us to summarize that in general, half-star ratings are less common than whole-star ratings (e.g., there are fewer ratings of 3.5 than there are ratings of 3 or 4, etc.):
```{r}
print(edx |> group_by(rating) |> summarize(count = n()))
```

We can visually see that from the following plot:
```{r}
edx |>
  group_by(rating) |>
  summarize(count = n()) |>
  ggplot(aes(x = rating, y = count)) +
  geom_line()

```

Further analysis of the `edx` dataset have been also inspired by the article mentioned above[@MRS-R-BEST], from which the code and explanatory notes below were cited.

##### Rating distribution plot[@MRS-R-BEST]
\
The code below demonstrates another way of visualizing the rating distribution:
```{r}
edx |>
  group_by(rating) |>
  summarize(count = n()) |>
  ggplot(aes(x = rating, y = count)) +
  geom_bar(stat = "identity", fill = "#8888ff") +
  ggtitle("Rating Distribution") +
  xlab("Rating") +
  ylab("Occurrences Count") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(n.breaks = 10) +
  theme_economist() +
  theme(axis.title.x = element_text(vjust = -5, face = "bold"), 
        axis.title.y = element_text(vjust = 10, face = "bold"), 
        plot.margin = margin(0.7, 0.5, 1, 1.2, "cm"))

```

This graph is another confirmation of what we found out above: rounded ratings occur more often than half-stared ones. The upward trend previously discussed is now perfectly clear, although it seems to top right between the 3 and 4-star ratings lowering the occurrences count afterward. That might be due to users being more hesitant to rate with the highest mark for whichever reasons they might hold[@MRS-R-BEST].

##### Yearly rating count[@MRS-R-BEST]
\
```{r}
print(edx |> 
  mutate(year = year(as_datetime(timestamp, origin = "1970-01-01"))) |>
  group_by(year) |>
  summarize(count = n())
)
```

##### Average rating per year plot[@MRS-R-BEST]
\
```{r}
edx |> 
  mutate(year = year(as_datetime(timestamp, origin = "1970-01-01"))) |>
  group_by(year) |>
  summarize(avg = mean(rating)) |>
  ggplot(aes(x = year, y = avg)) +
  geom_bar(stat = "identity", fill = "#8888ff") + 
  ggtitle("Average rating per year") +
  xlab("Year") +
  ylab("Average rating") +
  scale_y_continuous(labels = comma) + 
  theme_economist() +
  theme(axis.title.x = element_text(vjust = -5, face = "bold"), 
        axis.title.y = element_text(vjust = 10, face = "bold"), 
        plot.margin = margin(0.7, 0.5, 1, 1.2, "cm"))
```

##### Ratings per movie
\

###### Movie popularity count[@MRS-R-BEST]
\
```{r}
print(edx |> 
  group_by(movieId) |> 
  summarize(count = n()) |>
  slice_head(n = 10)
)
```

```{r}
summary(edx |> group_by(movieId) |> summarize(count = n()) |> select(count))
```

###### Ratings per movie plot[@MRS-R-BEST]
\
```{r}
edx |>
  group_by(movieId) |>
  summarize(count = n()) |>
  ggplot(aes(x = movieId, y = count)) +
  geom_point(alpha = 0.2, color = "#4020dd") +
  geom_smooth(color = "red") +
  ggtitle("Ratings per movie") +
  xlab("Movies") +
  ylab("Number of ratings") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(n.breaks = 10) +
  theme_economist() +
  theme(axis.title.x = element_text(vjust = -5, face = "bold"), 
        axis.title.y = element_text(vjust = 10, face = "bold"), 
        plot.margin = margin(0.7, 0.5, 1, 1.2, "cm"))
```

###### Movies' rating histogram[@MRS-R-BEST]
\
```{r}
edx |>
  group_by(movieId) |>
  summarize(count = n()) |>
  ggplot(aes(x = count)) +
  geom_histogram(fill = "#8888ff", color = "#4020dd") +
  ggtitle("Movies' rating histogram") +
  xlab("Rating count") +
  ylab("Number of movies") +
  scale_y_continuous(labels = comma) +
  scale_x_log10(n.breaks = 10) +
  theme_economist() +
  theme(axis.title.x = element_text(vjust = -5, face = "bold"), 
        axis.title.y = element_text(vjust = 10, face = "bold"), 
        plot.margin = margin(0.7, 0.5, 1, 1.2, "cm"))
```

##### Ratings per user[@MRS-R-BEST]
\

###### User rating count (activity measure)
\
```{r}
print(edx |> 
  group_by(userId) |> 
  summarize(count = n()) |>
  slice_head(n = 10)
)
```

###### User rating summary
\
```{r}
summary(edx |> group_by(userId) |> summarize(count = n()) |> select(count))
```

###### Ratings per user plot
\
```{r}
edx |>
  group_by(userId) |>
  summarize(count = n()) |>
  ggplot(aes(x = userId, y = count)) +
  geom_point(alpha = 0.2, color = "#4020dd") +
  geom_smooth(color = "red") +
  ggtitle("Ratings per user") +
  xlab("Users") +
  ylab("Number of ratings") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(n.breaks = 10) +
  theme_economist() +
  theme(axis.title.x = element_text(vjust = -5, face = "bold"), 
        axis.title.y = element_text(vjust = 10, face = "bold"), 
        plot.margin = margin(0.7, 0.5, 1, 1.2, "cm"))
```

###### Users' rating histogram
\
```{r}
edx |>
  group_by(userId) |>
  summarize(count = n()) |>
  ggplot(aes(x = count)) +
  geom_histogram(fill = "#8888ff", color = "#4020dd") +
  ggtitle("Users' rating histogram") +
  xlab("Rating count") +
  ylab("Number of users") +
  scale_y_continuous(labels = comma) +
  scale_x_log10(n.breaks = 10) +
  theme_economist() +
  theme(axis.title.x = element_text(vjust = -5, face = "bold"), 
        axis.title.y = element_text(vjust = 10, face = "bold"), 
        plot.margin = margin(0.7, 0.5, 1, 1.2, "cm"))
```

##### 
\
```{r}

```

##### 
\
```{r}

```

\
```{r echo=FALSE}
# #### [@MRS-R-BEST]
```

## Methods / Analysis
### Defining helper functions
\

Let's define some helper functions that we will use in our subsequent analysis:
```{r}
start_date <- function(){
  print(date())
  Sys.time()
}
end_date <- function(start){
  print(date())
  Sys.time() - start
}
rmse <- function(r) sqrt(mean(r^2))
RMSE <- function(true_ratings, predicted_ratings){
  sqrt(mean((true_ratings - predicted_ratings)^2))
}

# Because we know ratings can’t be below 0.5 or above 5, 
# we define the function clamp:
clamp <- function(x, min = 0.5, max = 5) pmax(pmin(x, max), min)

```

### Preparing train and set datasets
\
First, let's note that we have 10677 different movies: 
```{r}
n_movies <- n_distinct(edx$movieId)
print(n_movies)
```
and 69878 different users in the dataset:
```{r}
n_users <- n_distinct(edx$userId)
print(n_users)
```

Now, note the expressions below which confirm the fact explained in [Section _23.1.1 Movielens data_](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#movielens-data) of the _Course Textbook_[@IDS2] that not every user rated every movie:
```{r}
max_possible_ratings <- n_movies*n_users
sprintf("Maximum possible ratings: %s", max_possible_ratings)
sprintf("Rows in `edx` dataset: %s", dim_edx[1])
sprintf("Not every movie was rated: %s", max_possible_ratings > dim_edx[1])

```

As also explained in that section, we can think of these data as a very large matrix, with users on the rows and movies on the columns, with many empty cells. Therefore, we can think of a recommendation system as filling in the `NA`s in the dataset for the movies that some or all the users do not rate. A sample from the `edx` data below illustrates this idea[@IDS2_23-1-1]: 
```{r}
keep <- edx |> 
  dplyr::count(movieId) |> 
  top_n(4, n) |> 
  pull(movieId)

tab <- edx |> 
  filter(movieId %in% keep) |> 
  filter(userId %in% c(13:20)) |> 
  select(userId, title, rating) |> 
  mutate(title = str_remove(title, ", The"),
         title = str_remove(title, ":.*")) |>
  pivot_wider(names_from = "title", values_from = "rating")

print(tab)
```

The following plot of the matrix for a random sample of 100 movies and 100 users with yellow indicating a user/movie combination for which we have a rating shows how _sparse_ the matrix is:
```{r sparsity-of-movie-recs, echo=TRUE, fig.width=3, fig.height=3, out.width="40%"}
users <- sample(unique(edx$userId), 100)

rafalib::mypar()
edx|> 
  filter(userId %in% users) |> 
  select(userId, movieId, rating) |>
  mutate(rating = 1) |>
  pivot_wider(names_from = movieId, values_from = rating) |> 
  (\(mat) mat[, sample(ncol(mat), 100)])() |>
  as.matrix() |> 
  t() |>
  image(1:100, 1:100, z = _ , xlab = "Movies", ylab = "Users")
```

Further observations highlighted there that, as we can see from the distributions the author presented, some movies get rated more than others, and some users are more active than others in rating movies:
```{r movie-id-and-user-hists, echo=FALSE, fig.width=6, fig.height=3}
p1 <- edx |> 
  count(movieId) |> 
  ggplot(aes(n)) + 
  geom_histogram(bins = 30, color = "black") + 
  scale_x_log10() + 
  ggtitle("Movies")

p2 <- edx |> 
  count(userId) |> 
  ggplot(aes(n)) + 
  geom_histogram(bins = 30, color = "black") + 
  scale_x_log10() + 
  ggtitle("Users")

gridExtra::grid.arrange(p2, p1, ncol = 2)
```

Taking into consideration these observations, we came up with a decision to use the data from the dataset only for users who have provided at least 100 ratings.

Now let's split the `edx` dataset into a training set, which we will use to build and train our models, and a test set in which we will compute the accuracy of our predictions, the way described in the [Section 23.1.1 Movielens data](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#movielens-data) of the _Course Textbook_ mentioned above[@IDS2_23-1-1]:
```{r}
# Let's ignore the data for users who have not provided at least 100 ratings:
edx100 <- edx |> 
  group_by(userId) |>
  filter(n() >= 100) |>
  ungroup()

print(edx100 |> summarize(n_distinct(userId), n_distinct(movieId)))

# For each one of these users, we will split their ratings into 80% for training 
# and 20% for testing:

set.seed(2006)
indexes <- split(1:nrow(edx100), edx100$userId)
test_ind <- sapply(indexes, function(i) sample(i, ceiling(length(i)*.2))) |> 
  unlist() |>
  sort()

test_set <- edx100[test_ind,] 
train_set <- edx100[-test_ind,]

# To make sure we don’t include movies in the training set that should not be 
# there, we remove entries using the semi_join function:
test_set <- test_set |> semi_join(train_set, by = "movieId") |> as.data.frame()
summary(test_set)

train_set <- mutate(train_set, userId = factor(userId), movieId = factor(movieId))
summary(train_set)
```

We will use the array representation described in [Section 17.5 of the Textbook](https://rafalab.dfci.harvard.edu/dsbook-part-2/linear-models/treatment-effect-models.html#sec-anova), for the training data: we denote ranking for movie $j$ by user $i$ as $y_{i,j}$. To create this matrix, we use `tidyr::pivot_wider` function:

```{r}
y <- dplyr::select(train_set, movieId, userId, rating) |>
pivot_wider(names_from = movieId, values_from = rating) |>
column_to_rownames("userId") |>
as.matrix()

dim(y)
```

To be able to map movie IDs to titles we create the following lookup table:
```{r}
movie_map <- train_set |> dplyr::select(movieId, title, genres) |> 
  distinct(movieId, .keep_all = TRUE)

summary(movie_map)
```

Note that titles cannot be considered unique, so we can't use them as IDs[@IDS2_23-1-1].

### Naive Model
\
Let's begin our analysis by evaluating the simplest model described in [Section _23.3 The First Model_ of the *Course Textbook*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#a-first-model), and then gradually refine it through further research.
It is about a model that assumes the same rating for all movies and users with all the differences explained by random variation would look as follows:

$$
Y_{i,j} = \mu + \varepsilon_{i,j}
$$

with $\varepsilon_{i,j}$ independent errors sampled from the same distribution centered at 0 and $\mu$ the _true_ rating for all movies.

We know that the estimate that minimizes the RMSE is the least squares estimate of $\mu$ and, in this case, is the average of all ratings:

```{r pressure, echo=TRUE}
mu <- mean(y, na.rm = TRUE)
print(mu)
```

If we predict all unknown ratings with $\hat{\mu}$, we obtain the following RMSE:

```{r}
rmse(test_set$rating - mu)
```

If we plug in any other number, we will get a higher RMSE. Let's prove that by the following small investigation:
```{r}
deviation <- seq(0, 6, 0.1) - 3
print(deviation)

rmse_value <- sapply(deviation, function(diff){
  rmse(test_set$rating - mu + diff)
})

plot(deviation, rmse_value, type = "l")

sprintf("Minimum RMSE is achieved when the deviation from the mean is: %s", 
        deviation[which.min(rmse_value)])

```

To win the grand prize of $1,000,000, a participating team had to get an RMSE of at least 0.8563[@BigChaosSln]. So we can definitely do better![@IDS2_23-3]

### Taking into account User effects
\
To improve our model let's now take into consideration user effects as explained in [Section _23.4 User effects_ of the *Course Textbook*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#user-effects). 
If we visualize the average rating for each user the way the [the author](https://x.com/rafalab) shows, we can see that there is substantial variability in the average ratings across users: 
```{r}
hist(rowMeans(y, na.rm = TRUE), nclass = 30)
```

Following the author's further explanation, to account for this variability, we will use a linear model with a _treatment effect_  $\alpha_i$ for each user. The sum $\mu+\alpha_i$ can be interpreted as the typical rating user $i$ gives to movies. So we write the model as follows:

$$
Y_{i,j} = \mu + \alpha_i + \varepsilon_{i,j}
$$

Statistics textbooks refer to the $\alpha$s as treatment effects. In the Netflix challenge papers, they refer to them as _bias_[@IDS2_23-4; @MFT_RS].

As it is stated here[@IDS2_23-4], it can be shown that the least squares estimate $\hat{\alpha}_i$ is just the average of $y_{i,j} - \hat{\mu}$ for each user $i$. So we compute them this way:
```{r}
a <- rowMeans(y - mu, na.rm = TRUE)
```

Finally, we are ready to compute the `RMSE` (additionally using the helper function `clamp` we defined above to keep predictions in the proper range):
```{r}
# Compute the RMSE taking into account user effects:
user_effects_rmse <- test_set |> 
    left_join(data.frame(userId = as.integer(names(a)), a = a), by = "userId") |>
    mutate(resid = rating - clamp(mu + a)) |> 
    filter(!is.na(resid)) |>
    pull(resid) |> rmse()

print(user_effects_rmse)
```
### Taking into account Movie effects
\
In [Section _23.5 Movie effects_ of the *Course Textbook*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#movie-effects) the author draws our attention to the fact that some movies are generally rated higher than others. He also explains that a linear model with a _treatment effect_ $\beta_j$ for each movie can be used in this case, which can be interpreted as movie effect or the difference between the average ranking for movie $j$ and the overall average $\mu$:

$$
Y_{i,j} = \mu + \alpha_i + \beta_j +\varepsilon_{i,j}
$$
The author then shows how to use an approximation by first computing the least square estimate $\hat{\mu}$ and $\hat{\alpha}_i$, and then estimating $\hat{\beta}_j$ as the average of the residuals $y_{i,j} - \hat{\mu} - \hat{\alpha}_i$:

```{r}
b <- colMeans(y - mu - a, na.rm = TRUE)
```

We can now construct predictors and see how much the `RMSE` improves[@IDS2_23-5]:
```{r}
user_and_movie_effects_rmse <- test_set |> 
    left_join(data.frame(userId = as.integer(names(a)), a = a), by = "userId") |>
    left_join(data.frame(movieId = as.integer(names(b)), b = b), by = "movieId") |>
    mutate(resid = rating - clamp(mu + a + b)) |>  
    filter(!is.na(resid)) |>
    pull(resid) |> rmse()

print(user_and_movie_effects_rmse)
```
### Utilizing Penalized least squares
\
[Section _23.6 Penalized least squares_ of the *Course Textbook*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) explains why and how we should use _Penalized least squares_ to improve our predictions. The author also explains that the general idea of penalized regression is to control the total variability of the movie effects: $\sum_{j=1}^n \beta_j^2$. Specifically, instead of minimizing the least squares equation, we minimize an equation that adds a penalty:

$$ 
\sum_{i,j} \left(y_{u,i} - \mu - \alpha_i - \beta_j \right)^2 + \lambda \sum_{j} \beta_j^2
$$
The first term is just the sum of squares and the second is a penalty that gets larger when many $\beta_i$s are large. Using calculus, we can actually show that the values of $\beta_i$ that minimize this equation are:

$$
\hat{\beta}_j(\lambda) = \frac{1}{\lambda + n_j} \sum_{i=1}^{n_i} \left(Y_{i,j} - \mu - \alpha_i\right)
$$

where $n_j$ is the number of ratings made for movie $j$.

This approach will have our desired effect: when our sample size $n_j$ is very large, we obtain a stable estimate and the penalty $\lambda$ is effectively ignored since $n_j+\lambda \approx n_j$. Yet when the $n_j$ is small, then the estimate $\hat{\beta}_i(\lambda)$ is shrunken towards 0. The larger the $\lambda$, the more we shrink[@IDS2_23-6].

```{r}

```

## Conclusion

Hello Conclusion!

This is a great conclusion, isn't it?!!