### 2025-05-15 reports/capstone-movielens.site/02.8-1.MF.UMGYDE-model.Rmd
RMD report: `mean_reg` function described.

### UMGYDE Model: Matrix Factorization
\

\newpage