### 2025-05-20 reports/templates/02.X.UM+model.template.Rmd
Model RMD Template added.

## UM? Effect (UM?E) Model
\

### Mathematical Description of the UM?E Model
\

$$
Y_{i,j} = \mu + \alpha_i + \beta_j +\varepsilon_{i,j}
$$

If we define $g_{i,j}$ as the genre for user's $i$ rating of movie $j$, we can use the following models to account for the `genre` effect:
$$
Y_{i,j} = \mu + \alpha_i + \beta_j + g_{i,j} + \varepsilon_{i,j}
$$

where $g_{i,j}$ is an _aggregation function_ which is explained in detail in _Section 22.3: "Review of Aggregation Functions" of "Recommender Systems Handbook"_ (_Chapter 22: "Aggregation of Preferences in Recommender Systems"_, p. 712) book[@RRSK_RS_HB].

In the formula above $g_{i,j}$ denotes a _genre effect_ for user's $i$ rating of movie $j$, so that:

$$
g_{i,j} = \sum_{k=1}^K x_{i,j}^k \gamma_k
$$

with $x^k_{i,j} = 1$ if $g_{i,j}$ includes genre $k$, and $x^k_{i,j} = 0$ otherwise.

Therefore, for our current model, we can compute a predicted value 
$$
\hat{g}_{i,j} = g_{i,j} + \varepsilon_{i,j}
$$
as a residual: 
$$
\hat{g}_{i,j} = Y_{i,j} - (\mu + \alpha_i + \beta_j)
$$

### UM?E Model: Support Functions
\

::: {.noteblock data-latex=""}
The complete source code of the functions described in this section is available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/) section of the [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/) script on _GitHub_.
:::

#### [train_UM?_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L2) Function {#func.train_UM?_effect}
\

We use this function to build and train our model using the `train_set` dataset:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
The function described above accepts the `lambda` parameter, which we will need later for the _Model Regularization_ method. We also use the [mean_reg](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L63) function call, which we will need later for the *Regularization* techniques (for details, see the  [mean_reg](#func.mean_reg) function description in the Section [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report).
We have already explained that in the [UM? Effect Model Regularization] section. 
For now, we omit the `lambda` parameter, accepting its default value `lambda = 0`. Let's recall that in this case, the [mean_reg](#func.mean_reg) function is equivalent to the standard R function [base::mean](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/mean).
:::

#### `train_UM?E_effect.cv` Function {#func.train_UM?E_effect.cv}
\

We use the [train_UM?E_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/#L59) function to build and train our model using the `5-Fold Cross Validation` method. Below, we provide a slightly simplified version of the source code of that function:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
Here we use the function call [union_cv_results](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L432) to aggregate the _5-Fold Cross Validation_ method results (for details, see the [union_cv_results](#func.union_cv_results) function description in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report).
:::

#### `calc_UM?E_effect_MSE` Function
\

The source code of the function [calc_UM?E_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/#L) defined in the [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/) script to calculate the _Mean Squared Error (MSE)_ of the _UM?E Model_ for the given _Test Set_ is provided below:
```{r eval=FALSE}
```

\newpage

#### `calc_UM?E_effect_MSE.cv` Function
\

The source code of the function [calc_UM?E_effect_MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/#L) defined in the [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/) script to calculate the _5-Fold Cross Validation MSE_ result of the _UM?E Model_ is provided below:
```{r eval=FALSE}
```

#### `calc_UM?E_effect_RMSE` Function
\

The source code of the function [calc_UM?E_effect_RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/#L118) defined in the [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/) script to calculate the _Root Mean Squared Error (RMSE)_ of the _UM?E Model_ for the given _Test Set_ is provided below:

```{r eval=FALSE}
```

#### `calc_UM?E_effect_RMSE.cv` Function { #func.calc_UM?E_effect_RMSE.cv }
\

The source code of the function [calc_UM?E_effect_RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/#L) defined in the [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/) script to calculate the _5-Fold Cross Validation **RMSE**_ result of the _UM?E Model_ is provided below:

```{r eval=FALSE}
```

\newpage

### UM?E Model Building
\

::: {.noteblock data-latex=""}
The complete source code of builing and training the current model is available in the [Model building: UM? Effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R) script on _GitHub_.
:::

Below, we provide the most significant part of the code for training our model using the `5-Fold Cross Validation` method:
```{r, eval=FALSE}
```

\newpage

We can now construct predictors and calculate the _RMSE_ of the current model using the [calc_UM?E_effect_RMSE.cv](#func.calc_UM?E_effect_RMSE.cv) function described above:

```{r, eval=FALSE}
cv.UMG_effect.RMSE <- calc_UM?E_effect_RMSE.cv(cv.UMG_effect)

RMSEs.ResultTibble.UM?E <- RMSEs.ResultTibble.rglr.UM?E |> 
  RMSEs.AddRow("UM? Effect (UM?E) Model", cv.UMG_effect.RMSE,
               comment = "Comment")
```
```{r}
RMSE_kable(RMSEs.ResultTibble.UM?E)
```

::: {.noteblock data-latex=""}
Hello Note!!!
:::

\newpage

### UM? Effect Model Regularization
\

We have already explained the idea of the _Linear Model Regularization_ in the [Regularizing the User+Movie Effect Model] section above. Let's extend the concept outlined there to our current model.

$$
\sum_{j=1}^{n} \beta_j^2
$$ 
Specifically, instead of minimizing the least squares equation, we minimize an equation that adds a penalty:
$$ 
\sum_{i,j} \left(y_{u,i} - \mu - \alpha_i - \beta_j \right)^2 + \lambda \sum_{j} \beta_j^2
$$
The first term is just the sum of squares and the second is a penalty that gets larger when many $\beta_i$s are large. Using calculus, we can actually show that the values of $\beta_i$ that minimize this equation are:
$$
\hat{\beta}_j(\lambda) = \frac{1}{\lambda + n_j} \sum_{i=1}^{n_i} \left(Y_{i,j} - \mu - \alpha_i\right)
$$
where $n_j$ is the number of ratings made for movie $j$.

This approach will have our desired effect: when our sample size $n_j$ is very large, we obtain a stable estimate and the penalty $\lambda$ is effectively ignored since $n_j+\lambda \approx n_j$. Yet when the $n_j$ is small, then the estimate $\hat{\beta}_i(\lambda)$ is shrunken towards 0. The larger the $\lambda$, the more we shrink[@IDS2_23-6].

We will implement the _Regularization_ method on our  models (starting from the current model) in two steps:

  1. **Pre-configuration:** Preliminary determination of the optimal range of $\lambda$ values for the `5-Fold Cross Validation` samples;

  2. **Fine-tuning:** figuring out the value of $\lambda$ that minimizes the model's RMSE.

  3. **Retraining:** retraining the model with the best value of the parameter $\lambda$ obtained in the previous step.

\newpage

#### UM?E Model Regularization: Support Functions
\

::: {.noteblock data-latex=""}
The [regularize.test_lambda.UM?_effect.cv]() function described below are defined in the [Regularization]() section of the [UM-effect.functions.R]() script.
:::

##### [regularize.test_lambda.UM?_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L96) Function {#func.regularize.test_lambda.UM?_effect.cv}
\

This function calculates _RMSE_ of the _UM?E Model_ using _5-Fold Cross Validation_ method for the given $\lambda$ parameter value:
```{r, eval=FALSE}
```

::: {.noteblock data-latex=""}
Note that we reuse the function [train_UM?_effect.cv](#func.train_UM?_effect.cv) calling it from the [regularize.test_lambda.UM?_effect.cv](#func.regularize.test_lambda.UM?_effect.cv), but now with the $\lambda$ parameter different from the default ('lambda = 0') value.
:::

\newpage

#### UM?E Model Regularization: Pre-configuratioin
\

Let's perform the preconfiguration to determine the appropriate range of $\lambda$ for subsequent fine-tuning of our current model:

We are going to use the [tune.model_param](#func.tune.model_param) function described in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report, passing the [regularize.test_lambda.UM?_effect.cv](#func.regularize.test_lambda.UM?_effect.cv) function as the value of the [fn_tune.test.param_value](#func.tune.model_param.params.fn_tune.test.param_value) parameter:

```{r, eval=FALSE}
```

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section can be found in the [Preliminary setting-up of lambda range for UM? Effect regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L814) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R) script.
:::

Now, let's visualize the results of the $\lambda$ range preconfiguration:
``` {r}
```

::: {.noteblock data-latex=""}
We use the custom data visualization function [data.plot](#func.data.plot) described in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report, which is defined in the [Data Visualization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L447) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L592) script on _GitHub_.
:::

\newpage

#### UM?E Model Regularization: Fine-tuning
\

We are now ready to perform the fine-tuning step of our model _regularization_ process to determine the best value for the $\lambda$ parameter.

Here we are going to use the [model.tune.param_range](#func.model.tune.param_range) function described in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report, passing the [regularize.test_lambda.UM?_effect.cv](#func.regularize.test_lambda.UM?_effect.cv) function as the value of the [fn_tune.test.param_value](#func.model.tune.param_range.params.fn_tune.test.param_value) parameter:
```{r, eval=FALSE}
```

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section can be also found in the [Fine-tuning Step of the Regularization Method for the User+Movie Model](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L867) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R) script on _GitHub_.
:::

Let's visualize the fine-tuning results:
```{r}
```
\newpage

#### UM?E Model Regularization: Retraining Model with the best $\lambda$
\

Now, we can calculate the _Regularized UM? Effect_ by retraining our model on the entire `edx` dataset with the best value of the $\lambda$ parameter we just calculated, for the definitive _Root Mean Squared Error_ calculation and use in subsequent models.
```{r, eval=FALSE}
```

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section are available in the [Re-training Regularized UM? Effect Model for the best $\lambda$](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L910) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L910) script on _GitHub_.
:::

We calculate the _Root Mean Squared Error_ for the ultimately computed _UM? Effect_ using [calc_UM?_effect_RMSE.cv](#func.calc_UM?_effect_RMSE.cv) function described above as follows:
```{r, eval=FALSE}
```

Finally, we add the definitive result for the current model to our _Result Table_:
```{r, eval=FALSE}
RMSEs.ResultTibble.rglr.UM?E <- RMSEs.ResultTibble.UM?E |> 
  RMSEs.AddRow("Regularized UM? Effect Model", 
               UM?E.rglr.retrain.RMSE,
               comment = "Computed for `lambda` = %1" |>
                 msg.glue(UM?E.rglr.best_lambda))
```
```{r}
RMSE_kable(RMSEs.ResultTibble.rglr.UM?E)
```

\newpage