### 2025-05-23 reports/templates/02.X.UM+model.template.Rmd
UMGYE Model, Support functions described.

## UM? Effect (UM?E) Model
\

### Mathematical Description of the UM?E Model
\

If we define $\gamma(q)$ as the _year effect_ for the year (which is denotes by $q$ here) of rating a movie $j$ by a user $i$, the formula \@ref(eq:UMGE-model) describing the _UMGE Model_, for the current model, takes the following form:

\begin{equation}
Y_{i,j} = \mu + \alpha_i + \beta_j + g_{i,j} + \gamma(q) + \varepsilon_{i,j}
(\#eq:UM?E-model)
\end{equation}

Therefore, the formula \@ref(eq:genre-effect) for prediction a _genre effect_ as a residual, for a _year effect_ 
$$
\hat{\gamma}(q) = \gamma(q) + \varepsilon_{i,j}
$$

takes the following form:

\begin{equation}
\hat{\gamma}(q) = Y_{i,j} - (\mu + \alpha_i + \beta_j + g_{i,j})
(\#eq:year-effect)
\end{equation}

### UM?E Model: Support Functions
\

::: {.noteblock data-latex=""}
Some of the functions described in this section accepts the `lambda` parameter, which we will need later for the _Model Regularization_ method. We use the [mean_reg](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L63) function call as well, which we will also need for the *Regularization* techniques to apply to our models (for details, see the  [mean_reg](#func.mean_reg) function description in the Section [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report).
We have already explained that in the [UME Model Regularization] section. 
For now, we omit the `lambda` parameter, accepting its default value `lambda = 0`. Let's recall that in this case, the [mean_reg](#func.mean_reg) function is equivalent to the standard R function [base::mean](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/mean).
:::

#### [calc_UM?_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L70) Function {#func.calc_UM?_effect}
\

This is a helper function that calculates the _UM? effect_ of the user ratings and is intended to be used in other functions and scripts (for instance in [train_UM?_effect](#func.train_UM?_effect)) described below.

##### Parameters
\

The function accepts the only parameter `date_general_effect`, which is a data structure returned by any of the following functions described above:

  - [calc_date_general_effect](#func.calc_date_general_effect);
  - [calc_date_general_effect.cv](#func.calc_date_general_effect.cv).

##### Source Code
\

The following is the source code of the function:
```{r eval=FALSE}

```

::: {.noteblock data-latex=""}
The source code of the [calc_UM?_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L70) function is also available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L1) section of the [UM?-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L70) script on _GitHub_.
:::

##### Return

A data structure containg information about _User+Movie+Genre+Year (UM?) Effect_.

#### [train_UM?_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L75) Function {#func.train_UM?_effect}
\

This function is used to train the current model using the _Train Set_ dataset passed in the `train_set` parameter.
The following is the source code of the function:
```{r eval=FALSE}

```

::: {.noteblock data-latex=""}
The source code of the [train_UM?_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L75) function is also available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L1) section of the [UM?-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L75) script on _GitHub_.
:::

##### Return

A data structure containg information about _User+Movie+Genre+Year (UM?) Effect_.

#### [train_UM?_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L85) Function {#func.train_UM?_effect.cv}
\

This function is used to train the current model using the _5-Fold Cross Validation_ method.
The following is the source code of the function:
```{r eval=FALSE}

```

::: {.noteblock data-latex=""}
The source code of the [train_UM?_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L85) function is also available in the [UM?E Model Support Functions](#L1) section of the [UM?-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L85) script on _GitHub_.
:::

##### Return

A data structure containg information about _User+Movie+Genre+Year (UM?) Effect_.

#### [calc_UM?_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L93) Function
\

The function calculates the _Mean Squared Error (MSE)_ of the _UM?E Model_ for the given _Test Set_ passed in the `test_set` parameter.

##### Parameters
\

  - **test_set:** A _Test Set_ dataset used for _UM?E Model_ validation;
  - **UM?_effect:** A data structure returned by any of the following functions described above:
    - [calc_UM?_effect](#func.train_UM?_effect);
    - [train_UM?_effect](#train_UM?_effect) functions.
    - [train_UM?_effect.cv](#train_UM?_effect.cv).

##### Source Code
\

The source code of the function is provided below:
```{r eval=FALSE}

```

::: {.noteblock data-latex=""}
The source code of the function [calc_UM?_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L93) is also available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L93) script. 
:::

\newpage

#### [calc_UM?_effect_MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L119) Function
\

The function calculates the _5-Fold Cross Validation MSE_ result for the _UM? Effect_ data passed in the `UM?_effect` parameter.

##### Parameters
\

The function accepts the only parameter: `UM?_effect`, which is a data structure returned by the one of the following functions described above:

  - [calc_UM?_effect](#func.train_UM?_effect);
  - [train_UM?_effect](#train_UM?_effect) functions.
  - [train_UM?_effect.cv](#train_UM?_effect.cv).

##### Source Code
\

The source code of the function is provided below:
```{r eval=FALSE}
calc_UM?_effect_MSE.cv <- function(UM?_effect){
  start <- put_start_date()
  UM?_effect_MSEs <- sapply(edx_CV, function(cv_fold_dat){
    cv_fold_dat$validation_set |> calc_UM?_effect_MSE(UM?_effect)
  })
  put_end_date(start)
  put_log1("Function: calc_UM?_effect_MSE.cv
Date (Year) Effect MSE values have been computed for the %1-Fold Cross Validation samples.", 
           CVFolds_N)

  mean(UM?_effect_MSEs)
}

```

::: {.noteblock data-latex=""}
The source code of the function [calc_UM?_effect_MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L119) is also available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L119) script. 
:::

#### [calc_UM?_effect_RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L115) Function
\

The function calculates the _Mean Squared Error (MSE)_ of the _UM?E Model_ for the given _Test Set_ passed in the `test_set` parameter.

##### Parameters
\

  - **test_set:** A _Test Set_ dataset used for _UM?E Model_ validation;
  - **UM?_effect:** A data structure returned by any of the following functions described above:
    - [calc_UM?_effect](#func.train_UM?_effect);
    - [train_UM?_effect](#train_UM?_effect) functions.
    - [train_UM?_effect.cv](#train_UM?_effect.cv).

##### Source Code
\

The source code of the function is provided below:
```{r eval=FALSE}

```

::: {.noteblock data-latex=""}
The source code of the function [calc_UM?_effect_RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L115) is also available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L115) script. 
:::

\newpage

#### [calc_UM?_effect_RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L119) Function
\

The function calculates the _5-Fold Cross Validation MSE_ result for the _UM? Effect_ data passed in the `UM?_effect` parameter.

##### Parameters
\

The function accepts the only parameter: `UM?_effect`, which is a data structure returned by the one of the following functions described above:

  - [calc_UM?_effect](#func.train_UM?_effect);
  - [train_UM?_effect](#train_UM?_effect) functions.
  - [train_UM?_effect.cv](#train_UM?_effect.cv).

##### Source Code
\

The source code of the function is provided below:
```{r eval=FALSE}

```

::: {.noteblock data-latex=""}
The source code of the function [calc_UM?_effect_RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L119) is also available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L119) script. 
:::

\newpage

### UM?E Model Building
\

::: {.noteblock data-latex=""}
The complete source code of builing and training the current model is available in the [Model building: UM? Effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R) script on _GitHub_.
:::

Below, we provide the most significant part of the code for training our model using the `5-Fold Cross Validation` method:
```{r, eval=FALSE}
```

\newpage

We can now construct predictors and calculate the _RMSE_ of the current model using the [calc_UM?E_effect_RMSE.cv](#func.calc_UM?E_effect_RMSE.cv) function described above:

```{r, eval=FALSE}
cv.UMG_effect.RMSE <- calc_UM?E_effect_RMSE.cv(cv.UMG_effect)

RMSEs.ResultTibble.UM?E <- RMSEs.ResultTibble.rglr.UM?E |> 
  RMSEs.AddRow("UM? Effect (UM?E) Model", cv.UMG_effect.RMSE,
               comment = "Comment")
```
```{r}
RMSE_kable(RMSEs.ResultTibble.UM?E)
```

::: {.noteblock data-latex=""}
Hello Note!!!
:::

\newpage

### UM? Effect Model Regularization
\

We have already explained the idea of the _Linear Model Regularization_ in the [UME Model Regularization] section above. Let's extend the concept outlined there to our current model.

In this case, the formula \@ref(eq:ME-penalty) for adding a penalty takes the following form:

\begin{equation}
\sum_{i,j} \left(y_{u,i} - \mu - \alpha_i - \beta_j - ...\right)^2 + \lambda \sum_{j} \beta_j^2 
(\#eq:?E-penalty)
\end{equation}

And the formula \@ref(eq:ME-regularized) for calculating the values of the _treatment effect_ that minimize the equation takes the following form:

\begin{equation}
\hat{\beta}_j(\lambda) = \frac{1}{\lambda + n_j} \sum_{i=1}^{n_i} \left(Y_{i,j} - \mu - \alpha_i\right)
(\#eq:?E-regularized)
\end{equation}

where $n_$ is the number of ratings made for movie $j$.

As stated in the [UME Model Regularization] section, we will implement the _Regularization_ method on our  current model in three steps:

  1. **Pre-configuration:** Preliminary determination of the optimal range of $\lambda$ values for the `5-Fold Cross Validation` samples;

  2. **Fine-tuning:** figuring out the value of $\lambda$ that minimizes the model's RMSE.

  3. **Retraining:** retraining the model with the best value of the parameter $\lambda$ obtained in the previous step.

\newpage

#### UM?E Model Regularization: Support Functions
\

::: {.noteblock data-latex=""}
The [regularize.test_lambda.UM?_effect.cv]() function described below are defined in the [Regularization]() section of the [UM-effect.functions.R]() script.
:::

##### [regularize.test_lambda.UM?_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L96) Function {#func.regularize.test_lambda.UM?_effect.cv}
\

This function calculates _RMSE_ of the _UM?E Model_ using _5-Fold Cross Validation_ method for the given $\lambda$ parameter value:
```{r, eval=FALSE}
```

::: {.noteblock data-latex=""}
Note that we reuse the function [train_UM?_effect.cv](#func.train_UM?_effect.cv) calling it from the [regularize.test_lambda.UM?_effect.cv](#func.regularize.test_lambda.UM?_effect.cv), but now with the $\lambda$ parameter different from the default ('lambda = 0') value.
:::

\newpage

#### UM?E Model Regularization: Pre-configuration
\

Let's perform the preconfiguration to determine the appropriate range of $\lambda$ for subsequent fine-tuning of our current model:

We are going to use the [tune.model_param](#func.tune.model_param) function described in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report, passing the [regularize.test_lambda.UM?_effect.cv](#func.regularize.test_lambda.UM?_effect.cv) function as the value of the [fn_tune.test.param_value](#func.tune.model_param.params.fn_tune.test.param_value) parameter:

```{r, eval=FALSE}
```
```{r}
str(cv.UM?E.preset.result)
cv.UM?E.preset.result$best_result
```

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section can be found in the [UM?E Model Regularization: Pre-configuration](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R) script.
:::

Now, let's visualize the results of the $\lambda$ range preconfiguration:
``` {r}
# title = TeX(r'[UMGE Model Regularization: $\lambda$ Range Pre-configuration]')
```

::: {.noteblock data-latex=""}
We use the custom data visualization function [data.plot](#func.data.plot) described in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report, which is defined in the [Data Visualization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L447) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L592) script on _GitHub_.
:::

\newpage

#### UM?E Model Regularization: Fine-tuning
\

We are now ready to perform the fine-tuning step of our model _regularization_ process to determine the best value for the $\lambda$ parameter.

Here we are going to use the [model.tune.param_range](#func.model.tune.param_range) function described in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report, passing the [regularize.test_lambda.UM?_effect.cv](#func.regularize.test_lambda.UM?_effect.cv) function as the value of the [fn_tune.test.param_value](#func.model.tune.param_range.params.fn_tune.test.param_value) parameter:
```{r, eval=FALSE}
```
```{r}
str(UM?E.rglr.fine_tune.results)
UM?E.rglr.fine_tune.results$best_result
```

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section can be also found in the [Fine-tuning Step of the Regularization Method for the User+Movie Model](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L867) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R) script on _GitHub_.
:::

Let's visualize the fine-tuning results:
```{r}
```
\newpage

#### UM?E Model Regularization: Retraining Model with the best $\lambda$
\

Now, we can calculate the _Regularized UM? Effect_ by retraining our model on the entire `edx` dataset with the best value of the $\lambda$ parameter we just calculated, for the definitive _Root Mean Squared Error_ calculation and use in subsequent models.
```{r, eval=FALSE}
```

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section are available in the [Re-training Regularized UM? Effect Model for the best $\lambda$](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L910) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L910) script on _GitHub_.
:::

We calculate the _Root Mean Squared Error_ for the ultimately computed _UM? Effect_ using [calc_UM?_effect_RMSE.cv](#func.calc_UM?_effect_RMSE.cv) function described above as follows:
```{r, eval=FALSE}
```

Finally, we add the definitive result for the current model to our _Result Table_:
```{r, eval=FALSE}
RMSEs.ResultTibble.rglr.UM?E <- RMSEs.ResultTibble.UM?E |> 
  RMSEs.AddRow("Regularized UM? Effect Model", 
               UM?E.rglr.retrain.RMSE,
               comment = "Computed for `lambda` = %1" |>
                 msg.glue(UM?E.rglr.best_lambda))
```
```{r}
RMSE_kable(RMSEs.ResultTibble.rglr.UM?E)
```

\newpage

## Additional Function Templates
\

#### [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L) Function {#func.}
\

```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
Note
:::

#### [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L) Function {#func.}
\

```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
Note
:::

#### [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L) Function {#func.}
\

```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
Note
:::

#### [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L) Function {#func.}
\

```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
Note
:::

#### [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L) Function {#func.}
\

```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
Note
:::

\newpage