### 2025-05-23 reports/capstone-movielens.site/02.8-1.MF.UMGYDE-model.Rmd

### UMGYDE Model: Matrix Factorization
\

We use the following models to account for the `date` effect:

$$
Y_{i,j} = \mu + \alpha_i + \beta_j + f(d_{i,j}) + \varepsilon_{i,j}
$$

\newpage