### 2025-05-27 reports/templates/02.8.UMGYDE-model-template.Rmd
Started working on new model: model template added.

## User+Movie+Genre+Year+Day Effect (UMGYDE) Model
\

### Day Effect Analysis

::: {.noteblock data-latex=""}
The _Year Effect_ visualization and analysis code in this section are cited from the article [Movie Recommendation System using R - BEST](https://www.kaggle.com/code/amirmotefaker/movie-recommendation-system-using-r-best/notebook) mentioned earlier in this report[@MRS-R-BEST].
:::

### Mathematical Description of the UMGYDE Model
\

If we define $d_{i,j}$ as the _number of days since the earliest record_ in the `edx` dataset for movie $j$ rated by user $i$, then the formula \@ref(eq:UMGYE-model) describing the _UMGYE Model_, for the current model, takes the form:

\begin{equation}
Y_{i,j} = \mu + \alpha_i + \beta_j + g_{i,j} + \gamma(r_{i,j}) + f(d_{i,j}) \varepsilon_{i,j}
(\#eq:UMGYDE-model)
\end{equation}

with $f$ a smooth function of $d_{i,j}$

Therefore, the formula \@ref(eq:year-effect) for calculation the prediction of a _year effect_ as a residual, for a _smoothed day effect_ 
$$
\hat{f} = f(d_{i,j}) + \varepsilon_{i,j}
$$

takes the following form:

\begin{equation}
\hat{f(d_{i,j})} = Y_{i,j} - (\mu + \alpha_i + \beta_j + g_{i,j} + \gamma(r_{i,j}))
(\#eq:smoothed-day-effect)
\end{equation}

\newpage

### UMGYDE Model: Support Functions
\

::: {.noteblock data-latex=""}
Some of the functions described in this section accepts the `lambda` parameter, which we will need later for the _Model Regularization_ method. We use the [mean_reg](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L63) function call as well, which we will also need for the *Regularization* techniques to apply to our models (for details, see the  [mean_reg](#func.mean_reg) function description in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report).
We have already explained that in the [UME Model Regularization] section. 
For now, we omit the `lambda` parameter, accepting its default value `lambda = 0`. Let's recall that in this case, the [mean_reg](#func.mean_reg) function is equivalent to the standard R function [base::mean](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/mean).
:::

#### [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) Function {#func.calc_date_general_effect}
\

This is a helper function that calculates the _date general effect_ of the user ratings and is intended to be used in other functions and scripts (for instance in [train_UMGYD_effect](#func.train_UMGYD_effect)) described below.

The following is a slightly simplified version of the function's source code:
```{r eval=FALSE}
calc_date_general_effect <- function(train_set, lambda = 0){
  if (is.na(lambda)) {
    stop("Function: calc_date_general_effect
`lambda` is `NA`")
  }

  if(lambda == 0) put_log("Function `calc_date_general_effect`:
Computing Date Global Effect for given Train Set data...")
  else put_log1("Function `calc_date_general_effect`:
Computing Date Global Effect for lambda: %1...",
                lambda)
  dg_effect <- train_set |> 
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    mutate(resid = rating - (mu + a + b + g)) |>
  group_by(days) |>
    summarise(de = mean_reg(resid, lambda), 
              year = mean(year))
  dg_effect
}
```

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) function is available in the [UMGYDE Model Support Functions](#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) script on _GitHub_.
:::

##### Return

A data structure containg information about _Date General Effect_.

#### [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L30) Function {#func.calc_date_general_effect.cv}
\

This is a helper function that calculates the _date general effect_ of the user ratings using the _5-Fold Cross Validation_ method and is intended to be used in other functions and scripts (for instance in [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv)) described below.

##### Souurce Code
\

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L30) function is available in the [UMGYDE Model Support Functions](#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L30) script on _GitHub_.
:::

The following is a simplified version of the function's source code:
```{r eval=FALSE}
calc_date_general_effect.cv <- function(lambda = 0){
  if (is.na(lambda)) {
    stop("Function: calc_date_general_effect.cv
`lambda` is `NA`")
  }

  date_general_effect_ls <- lapply(edx_CV,  function(cv_fold_dat){
    cv_fold_dat$train_set |> calc_date_general_effect(lambda)
  })
  str(date_general_effect_ls)

  date_general_effect_united <- union_cv_results(date_general_effect_ls)

  date_general_effect_united |>
    group_by(days) |>
    summarise(de = mean(de, na.rm = TRUE), year = mean(year, na.rm = TRUE))
}
```

::: {.noteblock data-latex=""}
Here we use the function call [union_cv_results](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L432) to aggregate the _5-Fold Cross Validation_ method results (for details, see the [union_cv_results](#func.union_cv_results) function description in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report).
:::

##### Return

A data structure containg information about _Date General Effect_ as result of _5-Fold Cross Validation_.

\newpage

#### [calc_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L69) Function {#func.calc_UMGYD_effect}
\

This is a helper function that calculates the _UMGYD effect_ of the user ratings and is intended to be used in other functions and scripts (for instance in [train_UMGYD_effect](#func.train_UMGYD_effect)) described below.

##### Parameters
\

The function accepts the only parameter `date_general_effect`, which is a data structure returned by any of the following functions described above:

  - [calc_date_general_effect](#func.calc_date_general_effect);
  - [calc_date_general_effect.cv](#func.calc_date_general_effect.cv).

##### Source Code
\

The following is the source code of the function:
```{r eval=FALSE}
calc_UMGYD_effect <- function(date_general_effect){
  date_general_effect |>
    group_by(year) |>
    summarise(ye = mean(de, na.rm = TRUE))
}
```

::: {.noteblock data-latex=""}
The source code of the [calc_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L69) function is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L69) script on _GitHub_.
:::

##### Return

A data structure containg information about _User+Movie+Genre+Year (UMGYD) Effect_.

#### [train_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L74) Function {#func.train_UMGYD_effect}
\

This function is used to train the current model using the _Train Set_ dataset passed in the `train_set` parameter.
The following is the source code of the function:
```{r eval=FALSE}
train_UMGYD_effect <- function(train_set, lambda = 0){
  if (is.na(lambda)) {
    stop("Function: train_UMGYD_effect
`lambda` is `NA`")
  }

  train_set |>
    calc_date_general_effect(lambda) |>
    calc_UMGYD_effect()
}

```

::: {.noteblock data-latex=""}
The source code of the [train_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L74) function is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L74) script on _GitHub_.
:::

##### Return

A data structure containg information about _User+Movie+Genre+Year (UMGYD) Effect_.

#### [train_UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L84) Function {#func.train_UMGYD_effect.cv}
\

This function is used to train the current model using the _5-Fold Cross Validation_.
The following is the source code of the function:
```{r eval=FALSE}
train_UMGYD_effect.cv <- function(lambda = 0){
  if (is.na(lambda)) {
    stop("Function: train_UMGYD_effect.cv
`lambda` is `NA`")
  }

  calc_date_general_effect.cv(lambda) |> calc_UMGYD_effect()  
}
```

::: {.noteblock data-latex=""}
The source code of the [train_UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L84) function is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L84) script on _GitHub_.
:::

##### Return

A data structure containg information about _User+Movie+Genre+Year (UMGYD) Effect_.

#### [calc_UMGYD_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L92) Function {#func.calc_UMGYD_effect_MSE}
\

The function calculates the _Mean Squared Error (MSE)_ of the _UMGYDE Model_ for the given _Test Set_ passed in the `test_set` parameter.

##### Parameters
\

  - **test_set:** A _Test Set_ dataset used for _UMGYDE Model_ validation;
  - **UMGYD_effect:** A data structure returned by any of the following functions described above:
    - [calc_UMGYD_effect](#func.train_UMGYD_effect);
    - [train_UMGYD_effect](#func.train_UMGYD_effect);
    - [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv).

\newpage

##### Source Code
\

The source code of the function is provided below:
```{r eval=FALSE}
calc_UMGYD_effect_MSE <- function(test_set, UMGYD_effect){
  test_set |>
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    left_join(UMGYD_effect, by='year') |>
    mutate(resid = rating - clamp(mu + a + b + g + ye)) |> 
    pull(resid) |> mse()
}
```

::: {.noteblock data-latex=""}
The source code of the function [calc_UMGYD_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L92) is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L92) script. 
:::

#### [calc_UMGYD_effect_MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L102) Function {#func.calc_UMGYD_effect_MSE.cv}
\

The function calculates the _5-Fold Cross Validation MSE_ result for the _UMGYD Effect_ data passed in the `UMGYD_effect` parameter.

##### Parameters
\

The function accepts the only parameter: `UMGYD_effect`, which is a data structure returned by the one of the following functions described above:

  - [calc_UMGYD_effect](#func.train_UMGYD_effect);
  - [train_UMGYD_effect](#func.train_UMGYD_effect);
  - [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv).

##### Source Code
\

The source code of the function is provided below:
```{r eval=FALSE}
calc_UMGYD_effect_MSE.cv <- function(UMGYD_effect){
  start <- put_start_date()
  UMGYD_effect_MSEs <- sapply(edx_CV, function(cv_fold_dat){
    cv_fold_dat$validation_set |> calc_UMGYD_effect_MSE(UMGYD_effect)
  })
  put_end_date(start)
  put_log1("Function: calc_UMGYD_effect_MSE.cv
Date (Year) Effect MSE values have been computed for the %1-Fold Cross Validation samples.", 
           CVFolds_N)

  mean(UMGYD_effect_MSEs)
}

```

::: {.noteblock data-latex=""}
The source code of the function [calc_UMGYD_effect_MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L102) is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L102) script. 
:::

#### [calc_UMGYD_effect_RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L114) Function {#func.calc_UMGYD_effect_RMSE}
\

The function calculates the _Mean Squared Error (MSE)_ of the _UMGYDE Model_ for the given _Test Set_ passed in the `test_set` parameter.

##### Parameters
\

  - **test_set:** A _Test Set_ dataset used for _UMGYDE Model_ validation;
  - **UMGYD_effect:** A data structure returned by any of the following functions described above:
    - [calc_UMGYD_effect](#func.train_UMGYD_effect);
    - [train_UMGYD_effect](#func.train_UMGYD_effect);
    - [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv).

##### Source Code
\

The source code of the function is provided below:
```{r eval=FALSE}
calc_UMGYD_effect_RMSE <- function(test_set, UMGYD_effect){
  mse <- test_set |> calc_UMGYD_effect_MSE(UMGYD_effect)
  sqrt(mse)
}
```

::: {.noteblock data-latex=""}
The source code of the function [calc_UMGYD_effect_RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L114) is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L114) script. 
:::

#### [calc_UMGYD_effect_RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L118) Function {#func.calc_UMGYD_effect_RMSE.cv}
\

The function calculates the _5-Fold Cross Validation MSE_ result for the _UMGYD Effect_ data passed in the `UMGYD_effect` parameter.

##### Parameters
\

The function accepts the only parameter: `UMGYD_effect`, which is a data structure returned by the one of the following functions described above:

  - [calc_UMGYD_effect](#func.train_UMGYD_effect);
  - [train_UMGYD_effect](#func.train_UMGYD_effect);
  - [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv).

##### Source Code
\

The source code of the function is provided below:
```{r eval=FALSE}
calc_UMGYD_effect_RMSE.cv <- function(UMGYD_effect){
  UMGYD_effect_RMSE <- sqrt(calc_UMGYD_effect_MSE.cv(UMGYD_effect))
  put_log2("%1-Fold Cross Validation ultimate RMSE: %2", 
           CVFolds_N, 
           UMGYD_effect_RMSE)

  UMGYD_effect_RMSE
}
```

::: {.noteblock data-latex=""}
The source code of the function [calc_UMGYD_effect_RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L118) is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L118) script. 
:::

\newpage

### UMGYDE Model Building
\

::: {.noteblock data-latex=""}
The complete source code of building and training the current model is available in the [UMGYDE Model Building](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1432) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1432) script on _GitHub_.
:::

Below, we provide the most significant line of the code for training our model using the `5-Fold Cross Validation` method:
```{r, eval=FALSE}
  cv.UMGYD_effect <- train_UMGYD_effect.cv()
```
```{r}
str(cv.UMGYD_effect)
summary(cv.UMGYD_effect)
```

::: {.noteblock data-latex=""}
We use the [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv) function described above to compute the _UMGYD Effect_.
:::

We can now construct predictors and calculate the _RMSE_ of the current model using the [calc_UMGYD_effect_RMSE.cv](#func.calc_UMGYD_effect_RMSE.cv) function described above:

```{r, eval=FALSE}
cv.UMGYD_effect.RMSE <- calc_UMGYD_effect_RMSE.cv(cv.UMGYD_effect)

RMSEs.ResultTibble.UMGYDE <- RMSEs.ResultTibble.rglr.UMGYE |> 
  RMSEs.AddRow("UMGYDE Model", 
               cv.UMGYD_effect.RMSE,
               comment = "User+Movie+Genre+Year+Day Effect (UMGYDE) Model")
```
```{r}
RMSE_kable(RMSEs.ResultTibble.UMGYDE)
```

::: {.noteblock data-latex=""}
Now, we have a bit better result than one for the previous model.
:::

\newpage

### UMGYDE Model Regularization
\

We have already explained the idea of _Linear Model Regularization_ in the [UME Model Regularization] section above. We have also seen how the formula \@ref(eq:ME-penalty) for adding a penalty to the _UME Model_ is transformed into the formula \@ref(eq:GE-penalty) for the _UMGE Model_ and then into the formula \@ref(eq:GYE-penalty) for the _UMGYE Model_. For the current model, this formula takes the form:

\begin{equation}
\sum_{i,j} \left(y_{u,i} - \mu - \alpha_i - \beta_j - g_{i,j} - \gamma(r_{i,j}) - f(d_{i,j})\right)^2 + \lambda \sum_{i,j} f(d_{i,j})^2
(\#eq:DE-penalty)
\end{equation}

And the formula \@ref(eq:ME-regularized) for calculating the values of the _treatment effect_ that minimizes the equation will take the form:

\begin{equation}
\hat{\gamma}(r, \lambda) = \frac{1}{\lambda + n_d} \sum_{i=1}^{n_{d}} \left(Y_{i,j} - \mu - \alpha_i - \beta_j - g_{i,j} - \gamma(r)\right)
(\#eq:DE-regularized)
\end{equation}

where $n_d$ is the number of ratings made on the day $d$.

As with the previous models, we implement the _Regularization_ method for the current model in the following three steps:

  1. **Pre-configuration:** Preliminary determination of the optimal range of $\lambda$ values for the `5-Fold Cross Validation` samples;

  2. **Fine-tuning:** figuring out the value of $\lambda$ that minimizes the model's RMSE.

  3. **Retraining:** retraining the model with the best value of the parameter $\lambda$ obtained in the previous step.

\newpage

#### UMGYDE Model Regularization: Support Functions
\

::: {.noteblock data-latex=""}
The [regularize.test_lambda.UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L128) function described below are defined in the [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L127) section of the [UM-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L128) script.
:::

##### [regularize.test_lambda.UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#128) Function {#func.regularize.test_lambda.UMGYD_effect.cv}
\

This function calculates _RMSE_ of the _UMGYDE Model_ using _5-Fold Cross Validation_ for the given $\lambda$ parameter value:
```{r, eval=FALSE}
regularize.test_lambda.UMGYD_effect.cv <- function(lambda){
  if (is.na(lambda)) {
    stop("Function: regularize.test_lambda.UMGYD_effect.cv
`lambda` is `NA`")
  }

  train_UMGYD_effect.cv(lambda) |>
    calc_UMGYD_effect_RMSE.cv()
}
```

::: {.noteblock data-latex=""}
Note that we reuse the function [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv) calling it from the [regularize.test_lambda.UMGYD_effect.cv](#func.regularize.test_lambda.UMGYD_effect.cv), but now with the $\lambda$ parameter different from the default ('lambda = 0') value.
:::

\newpage

#### UMGYDE Model Regularization: Pre-configuration
\

Let's perform the preconfiguration to determine the appropriate range of $\lambda$ for subsequent fine-tuning of our current model:

We are going to use the [tune.model_param](#func.tune.model_param) function described in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report, passing the [regularize.test_lambda.UMGYD_effect.cv](#func.regularize.test_lambda.UMGYD_effect.cv) function as the value of the [fn_tune.test.param_value](#func.tune.model_param.params.fn_tune.test.param_value) parameter.

Below we provide the most significant part of the code that performs this operation:
```{r, eval=FALSE}
  lambdas <- seq(0, 512, 32)
  cv.UMGYDE.preset.result <- 
    tune.model_param(lambdas, 
                     regularize.test_lambda.UMGYD_effect.cv,
                     steps.beyond_min = 16)
```
```{r}
str(cv.UMGYDE.preset.result)
cv.UMGYDE.preset.result$best_result
```

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section can be found in the [UMGYDE Model Regularization: Pre-configuration](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1493) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1493) script on _GitHub_.
:::

Now, let's visualize the results of the $\lambda$ range preconfiguration:
``` {r}
cv.UMGYDE.preset.result$tuned.result |>
  data.plot(title = TeX(r'[UMGYDE Model Regularization: $\lambda$ Range Pre-configuration]'),
              xname = "parameter.value", 
              yname = "RMSE", 
              xlabel = TeX(r'[$\lambda$]'), 
              ylabel = "RMSE")
```

::: {.noteblock data-latex=""}
We use the custom data visualization function [data.plot](#func.data.plot) described in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report, which is defined in the [Data Visualization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L447) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L592) script on _GitHub_.
:::

\newpage

#### UMGYDE Model Regularization: Fine-tuning
\

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section can be found in the [Fine-tuning Step of the Regularization Method for the User+Movie Model](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L867) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R) script on _GitHub_.
:::

We are now ready to perform the fine-tuning step of our model _regularization_ process to determine the best value for the $\lambda$ parameter.

Here we are going to use the [model.tune.param_range](#func.model.tune.param_range) function described in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report, passing the [regularize.test_lambda.UMGYD_effect.cv](#func.regularize.test_lambda.UMGYD_effect.cv) function as the value of the [fn_tune.test.param_value](#func.model.tune.param_range.params.fn_tune.test.param_value) parameter.

Below we provide the most significant part of the code that performs this operation:
```{r, eval=FALSE}
endpoints <- 
  get_fine_tune.param.endpoints(cv.UMGYDE.preset.result$tuned.result)

UMGYDE.loop_starter <- c(endpoints["start"], 
                        endpoints["end"], 
                        8)
UMGYDE.loop_starter
#> [1]

cache.base_name <- "UMGYDE.rglr.fine-tuning"

UMGYDE.rglr.fine_tune.results <- 
  model.tune.param_range(UMGYDE.loop_starter,
                         UMGYDE.rglr.fine_tune.cache.path,
                         cache.base_name,
                         regularize.test_lambda.UMGYD_effect.cv)
```
```{r}
str(UMGYDE.rglr.fine_tune.results)
UMGYDE.rglr.fine_tune.results$best_result
```

::: {.noteblock data-latex=""}
Under the hood, this code block internally calls the helper function [get_fine_tune.param.endpoints](#func.get_fine_tune.param.endpoints) described in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report to figure out the boundaries of the range of values from the nearest neighborhood of the value corresponding to the best `RMSE` in the data passed in as a parameter.
:::

\newpage

Let's visualize the fine-tuning results:
```{r}
UMGYDE.rglr.fine_tune.results$tuned.result |>
  data.plot(title = "UMGYDE Model Regularization: Fine-tuned result",
              xname = "parameter.value",
              yname = "RMSE",
              xlabel = TeX(r'[$\lambda$]'),
              ylabel = str_glue("Deviation from the best RMSE value (",
                                as.character(round(UMGYDE.rglr.fine_tune.RMSE.best, digits = 7)),
                                ")"),
              normalize = TRUE)
```

::: {.noteblock data-latex=""}
We use the custom data visualization function [data.plot](#func.data.plot) described in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report, which is defined in the [Data Visualization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L447) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L592) script on _GitHub_.
:::

\newpage

#### UMGYDE Model Regularization: Retraining Model with the best $\lambda$
\

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section are available in the [Re-train Regularized UMGYDE Model for the best `lambda`](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1589) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1589) script on _GitHub_.
:::

Now, we can calculate the _Regularized UMGYDE Effect_ by retraining our model on the entire `edx` dataset with the best value of the $\lambda$ parameter we just calculated, for the definitive _Root Mean Squared Error_ calculation and use in subsequent models.

```{r, eval=FALSE}
  rglr.UMGYD_effect <- edx |> train_UMGYD_effect(UMGYDE.rglr.best_lambda)

```

::: {.noteblock data-latex=""}
Here we use the helper function [train_UMGYD_effect](#func.train_UMGYD_effect) described above in the [UMGYDE Model Regularization: Support Functions] section.
:::

We calculate the _Root Mean Squared Error_ for the ultimately computed _UMGYDE Effect_ using the helper function [calc_UMGYD_effect_RMSE.cv](#func.calc_UMGYD_effect_RMSE.cv) described above in the [UMGYDE Model Regularization: Support Functions] section as follows:
```{r, eval=FALSE}
  rglr.UMGYD_effect.RMSE <- calc_UMGYD_effect_RMSE.cv(rglr.UMGYD_effect)
```

Finally, we add the definitive result for the current model to our _Result Table_:
```{r, eval=FALSE}
RMSEs.ResultTibble.rglr.UMGYDE <- RMSEs.ResultTibble.UMGYDE |> 
  RMSEs.AddRow("Regularized UMGYDE Model", 
               rglr.UMGYD_effect.RMSE,
               comment = "Computed for `lambda` = %1" |>
                 msg.glue(UMGYDE.rglr.best_lambda))

```
```{r}
RMSE_kable(RMSEs.ResultTibble.rglr.UMGYDE)
```

\newpage