### 2025-06-06 reports/capstone-movielens.site/03.conclusion.Rmd
Working on Conclution section (in progress)

# Conclusion
\

In this project, a _Machine Learning Algorithm_ was successfully developed for a _Recommender System_ (hereinafter referred to as _RS MLA_) to predict movie ratings using the [10M version of the MovieLens](http://grouplens.org/datasets/movielens/10m/) dataset.

Since it was needed to convert the original dataset into a _Data Source_, suitable for processing by the _RS MLA_, a package [edx.capstone.movielens.data], specifically designed for this purpose, was developed and used, providing the _RS MLA_ with the necessary datasets.

Following the Netflix Grand Prize Contest requirements, the _Root Mean Squared Error (RMSE)_ metric was used to evaluate the accuracy of our models.

By incorporating essential factors such as overall average rating, movie, user, and some other observable effects, as well as efficient techniques such as _Regularization_ and _Matrix Factorization_ methods, the accuracy was significantly improved compared to the baseline model.

The baseline model was the _Overall Mean Rating_ model described in the [Overall Mean Rating (OMR) Model] section.

The following effects were taken into account:

  - ***User:*** based on which the _User Effect (UE) Model_ described in the [User Effect (UE) Model] section was built;
  - ***Movie:*** ;
  - ***Genre:*** ;
  - ***Year:*** ;
  - ***Day*** ;.

\newpage