### 2025-09-18 reports/capstone-movielens.site/03.results.Rmd
Last update made on June 2025.

# Results
\

The *Result Summary Table* constructed below shows in detail the progress of our *Recommender System* improvement:

```{r, eval=FALSE}
total.RMSEs.ResultTibble <- RMSEs.AddDiffColumn(final.MF.RMSEs.ResultTibble)
```
```{r}
RMSE.Total_kable(total.RMSEs.ResultTibble)
```

::: {.noteblock data-latex=""}
Here use the [RMSEs.AddDiffColumn](#func.RMSEs.AddDiffColumn) and [RMSE.Total_kable](#func.RMSE.Total_kable) functions described in the [Common Helper Functions: Result RMSEs Table] section of the [Appendix] to this report.
:::

\newpage