### 2025-09-29 reports/capstone-movielens.site/work/draft-snippert.Rmd
Report: Fix for `clamp` function description.

---
title: "func-descr.template"
output: html_document
---

#### [clamp](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L6) Function {#func.clamp}
\

Clamps the input value passed in [r](#func.clamp.args.r) argument between the values passed in [min](#func.clamp.args.min) and [max](#func.clamp.args.max) arguments.

::: { #func.clamp.Usage .sidebar }
##### Usage
\

```{r eval=FALSE}
r_clamped <- clamp(r, min = 0.5, max = 5)
```
:::

::: { #func.clamp.args .sidebar }
##### Arguments
\

::: { #func.clamp.args.r .sidebar }
  - **r:** Value to clamp;

:::
::: { #func.clamp.args.min .sidebar }
  - **min:** Minimum allowed value;

:::
:::
::: { #func.clamp.args.max .sidebar }
  - **max:** Maximum allowed value;

:::
:::

::: { #func.clamp.details .sidebar }
##### Details
\

As explained in [Section  23.4 *User effects*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#user-effects) of the _Course Textbook_ we know ratings cannot be below 0.5 or above 5. For this reason, we need a function described in that section[@IDS2_23-4]:

Therefore, in our project this function is used to clamp the computed movie rating values between 0.5 and 5.

:::

::: { #func.clamp.value .sidebar }
##### Value
\

The value clamped between [min](#func.clamp.args.min) and [max](#func.clamp.args.max) argument values.

:::

::: { #func.clamp.code .sidebar }
##### Source Code
\

The source code of the [clamp](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L6) function is shown below:
```{r eval=FALSE}
clamp <- function(x, min = 0.5, max = 5) pmax(pmin(x, max), min)
```

::: {.noteblock data-latex=""}
The source code of the [clamp](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L6) function is also available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L3) section of the [common-helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L6) script on _GitHub_.
:::
:::