### 2025-09-29 reports/capstone-movielens.site/work/draft-snippert.Rmd
Capstone Report: `open_logfile` function description added.

---
title: "func-descr.template"
output: html_document
---

#### [open_logfile](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/logging-functions.R#L3) Function {#func.open_logfile}
\

A function to initialize the log file.

::: { #func.open_logfile.Usage .sidebar }
##### Usage
\

```{r eval=FALSE}
open_logfile(file_name)
```
:::

::: { #func.open_logfile.args .sidebar }
##### Arguments
\

::: { #func.open_logfile.args.r .sidebar }
  - **file_name:** A name used as a base to generate a unique log file name;

:::
:::

::: { #func.open_logfile.details .sidebar }
##### Details
\

This function is a wrapper to [logr::log_open](https://www.rdocumentation.org/packages/logr/versions/1.3.9/topics/log_open) function. It also calls internally the following functions to generate a unique file name based on the [file_name] argument and _system time_:
  - [Sys.time]();
  - [stringr::str_replace](https://www.rdocumentation.org/packages/stringr/versions/1.5.0/topics/str_replace);
  - [stringr::str_replace_all](https://www.rdocumentation.org/packages/stringr/versions/1.5.0/topics/str_replace)

For more information on the _inner functions_ listed above, see the [R Documentation](https://www.rdocumentation.org/).

:::

::: { #func.open_logfile.value .sidebar }
##### Value
\

The path of the log.

:::

::: { #func.open_logfile.code .sidebar }
##### Source Code
\

The source code of the [open_logfile](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/logging-functions.R#L3) function is shown below:
```{r eval=FALSE}
library(logr)
library(stringr)

open_logfile <- function(file_name){
  log_file_name <- as.character(Sys.time()) |> 
    str_replace_all(':', '_') |> 
    str_replace(' ', 'T') |>
    str_c(file_name)

  log_open(file_name = log_file_name)
}
```

::: {.noteblock data-latex=""}
The source code of the [open_logfile](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/logging-functions.R#L3) function is also available in the [Logging Helper functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/logging-functions.R#L1) section of the [logging-functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/logging-functions.R#L3) script on _GitHub_.
:::
:::