### 2025-11-06 reports/capstone-movielens.site/work/draft-snippert.Rmd

---
title: "func-descr.template"
output: html_document
---

::: {.noteblock data-latex=""}
In the code snippet above, we use the custom function [@fn](#func.@fn) described in the [sec_name] section of [Appendix A: Support Functions].
:::